---
title: "Data 621 - Assignment 1"
date: "June 19, 2016"
output: html_document
---

```{r ECHO=FALSE}
library(leaps) #not relevant, we need bestglm
library(bestglm)
library(ISLR)
library(dplyr)
library(ggplot2)



# To clear a Global Environemnt
rm(list =ls())
```

#Load the data
Our first step is to load all the required data, this includes the cleaned-up data.

```{r}
# Read in the data files 
file_name <- "~/Downloads/Data/crime-training-data.csv"
crime_data <- read.csv(file_name, header=TRUE, stringsAsFactors=FALSE, fileEncoding="latin1")

file_name <- "~/Downloads/Data/crime-evaluation-data.csv"
crime_eval_data <- read.csv(file_name, header=TRUE, stringsAsFactors=FALSE, fileEncoding="latin1")

```

Now that we have all the data, we should examine it.

```{r}
summary(crime_data)
#Let's take a look at the box plots for some of the values
boxplot(zn~target, data=crime_data, ylab="zn",xlab="Above median crime rate? (0=No, 1=Yes)")
boxplot(indus~target, data=crime_data, ylab="indus",xlab="Above median crime rate? (0=No, 1=Yes)")
boxplot(chas~target, data=crime_data, ylab="chas",xlab="Above median crime rate? (0=No, 1=Yes)")
boxplot(nox~target, data=crime_data, ylab="nox",xlab="Above median crime rate? (0=No, 1=Yes)")
boxplot(rm~target, data=crime_data, ylab="rm",xlab="Above median crime rate? (0=No, 1=Yes)")
boxplot(dis~target, data=crime_data, ylab="dis",xlab="Above median crime rate? (0=No, 1=Yes)")
boxplot(age~target, data=crime_data, ylab="age",xlab="Above median crime rate? (0=No, 1=Yes)")
boxplot(dis~target, data=crime_data, ylab="dis",xlab="Above median crime rate? (0=No, 1=Yes)")
boxplot(rad~target, data=crime_data, ylab="rad",xlab="Above median crime rate? (0=No, 1=Yes)")
boxplot(tax~target, data=crime_data, ylab="tax",xlab="Above median crime rate? (0=No, 1=Yes)")
boxplot(ptratio~target, data=crime_data, ylab="ptratio",xlab="Above median crime rate? (0=No, 1=Yes)")
boxplot(black~target, data=crime_data, ylab="black",xlab="Above median crime rate? (0=No, 1=Yes)")
boxplot(lstat~target, data=crime_data, ylab="lstat",xlab="Above median crime rate? (0=No, 1=Yes)")
boxplot(medv~target, data=crime_data, ylab="medv",xlab="Above median crime rate? (0=No, 1=Yes)")

ggplot(crime_data, aes(x = zn)) + geom_density() #right-skewed normal
ggplot(crime_data, aes(x = indus)) + geom_density() #NOT normal
ggplot(crime_data, aes(x = nox)) + geom_density() #Poisson
ggplot(crime_data, aes(x = rm)) + geom_density() #normal
ggplot(crime_data, aes(x = age)) + geom_density() #some other distribution, seems wrong
ggplot(crime_data, aes(x = dis)) + geom_density() #Poisson
ggplot(crime_data, aes(x = rad)) + geom_density() #Nor normal - two humps
ggplot(crime_data, aes(x = tax)) + geom_density() #Not normal (two humps)
ggplot(crime_data, aes(x = ptratio)) + geom_density() #severely left skewed

pairs(~zn+indus+nox+rm+age+dis+rad+tax+ptratio,data=crime_data,gap=0.4,cex.labels=1.5)


#convert chas into a fact
crime_data$chas <- factor(crime_data$chas)
crime_eval_data$chas <- factor(crime_eval_data$chas)
```
# Using the glm function
```{r}

glm.fit <- glm(target~.,data=crime_data,family=binomial(link="logit"))
summary(glm.fit)

glm.probs <-predict(glm.fit,type="response")
glm.probs[1:5]
glm.pred <- ifelse(glm.probs>0.5, "Above median crime rate", "Below median crime rate")
attach(crime_data)
table(glm.pred,target)
mean(glm.pred==target)

1-pchisq(186.1,452) #p-value of residuals
1-pchisq(645.9,465) #p-value of Null
#Results indicate this is a very strong model...
```

# Backwards model technique
```{r}
glm.fit2 <- update(glm.fit, ~ . - lstat)
summary(glm.fit2)
#This model is not superior given the lower AIC, we shall continue with the first

exp(glm.fit$coefficients) #Odds ratios

#marginal effects
LogitScalar <- mean(dlogis(predict(glm.fit, type = "link")))
LogitScalar * coef(glm.fit)

#predicted probabilities
plogit <- predict(glm.fit, type="response")
summary(plogit)

glm.fit3 <- update(glm.fit, ~ . - indus)
summary(glm.fit3)

#Let's remove all variables that are not significant
glm.fit4 <- glm(target~nox+age+dis+rad+tax+ptratio+black+medv, family = binomial(link = "logit"))
summary(glm.fit4)

#We take a look at confusion matrices to help get a sense of which model is best
table(true = target, pred = round(fitted(glm.fit)))
table(true = target, pred = round(fitted(glm.fit2)))
table(true = target, pred = round(fitted(glm.fit3)))
table(true = target, pred = round(fitted(glm.fit4)))

#McFadden Adjusted R-squared
logit0 <- update(glm.fit, formula=target~1)
McFadden <- 1-as.vector(logLik(glm.fit)/logLik(logit0))
McFadden

logit0 <- update(glm.fit2, formula=target~1)
McFadden <- 1-as.vector(logLik(glm.fit2)/logLik(logit0))
McFadden

logit0 <- update(glm.fit3, formula=target~1)
McFadden <- 1-as.vector(logLik(glm.fit3)/logLik(logit0))
McFadden

logit0 <- update(glm.fit4, formula=target~1)
McFadden <- 1-as.vector(logLik(glm.fit4)/logLik(logit0))
McFadden

confint(glm.fit)

#reviewing these, a change in sign indicates a really poor choice of coefficient, so we should remove zn, indus, rm, tax
#Let's remove all variables that are not significant
glm.fit5 <- glm(target~nox+age+dis+rad+ptratio+medv, family = binomial(link = "logit"))
summary(glm.fit5)

#Results, however are worse
```

#Further analysis of the glm function
```{r}
loessfit1 <- loess(target ~ nox,degree=1,span=2/3)
loessfit2 <- loess(glm.fit$fitted.values ~ nox,degree=1,span=2/3)
xx <- seq(.35,.9,0.05)
par(mfrow=c(1,2))
plot(nox,target,xlab="Fnox, x1", ylab="Y, Above median? (0=No, 1=Yes)")
lines(xx,predict(loessfit1,data.frame(nox=xx)))

plot(nox,glm.fit$fitted.values,ylab=expression(hat(Y)),xlab="nox, x1")
lines(xx,predict(loessfit2,data.frame(nox=xx)))
```


#Predicting the results (glm)
```{r}

#Prediction model
pm <- predict(glm.fit,crime_eval_data, type="response")
summary(pm)
```

#Trying the glmulti package
```{r}
#from https://rstudio-pubs-static.s3.amazonaws.com/2897_9220b21cfc0c43a396ff9abf122bb351.html
library(glmulti)
glmulti.logistic.out <-
    glmulti(target~., data = crime_data,
            level = 1,               # No interaction considered
            method = "h",            # Exhaustive approach
            crit = "aic",            # AIC as criteria
            confsetsize = 5,         # Keep 5 best models
            plotty = F, report = F,  # No plot or interim reports
            fitfunction = "glm",     # glm function
            family = binomial)       # binomial family for logistic regression

## Show 5 best models (Use @ instead of $ for an S4 object)
glmulti.logistic.out@formulas

## Show result for the best model
summary(glmulti.logistic.out@objects[[1]])

#Analysis of the model
gmu <- glmulti.logistic.out@objects[[1]]
table(true = target, pred = round(fitted(gmu)))

McFadden <- 1-as.vector(logLik(gmu)/logLik(logit0))
McFadden

#Try again with new criteria (bic)
glmulti.logistic.bic.out <-
    glmulti(target~., data = crime_data,
            level = 1,               # No interaction considered
            method = "h",            # Exhaustive approach
            crit = "bic",            # BIC as criteria
            confsetsize = 5,         # Keep 5 best models
            plotty = F, report = F,  # No plot or interim reports
            fitfunction = "glm",     # glm function
            family = binomial)       # binomial family for logistic regression

## Show 5 best models (Use @ instead of $ for an S4 object)
glmulti.logistic.bic.out@formulas

## Show result for the best model
summary(glmulti.logistic.bic.out@objects[[1]])
#Note that residual deviance is worse for BIC method vs. AIC

gmu.bic <- glmulti.logistic.bic.out@objects[[1]]
table(true = target, pred = round(fitted(gmu.bic)))

McFadden <- 1-as.vector(logLik(gmu)/logLik(logit0))
McFadden
```

#Further analysis of the glmulti function
```{r}
loessfit1 <- loess(target ~ nox,degree=1,span=2/3)
loessfit2 <- loess(gmu$fitted.values ~ nox,degree=1,span=2/3)
xx <- seq(.35,.9,0.05)
summary(gmu)
par(mfrow=c(1,2))
plot(nox,target,xlab="Fnox, x1", ylab="Y, Above median? (0=No, 1=Yes)")
lines(xx,predict(loessfit1,data.frame(nox=xx)))

plot(nox,gmu$fitted.values,ylab=expression(hat(Y)),xlab="nox, x1")
lines(xx,predict(loessfit2,data.frame(nox=xx)))
```

#Predicting the results (glmulti)
```{r}

#Prediction model
pm <- predict(gmu,crime_eval_data, type="response")
summary(pm)
```

# Using the leaps/bestglm package (ignore now, maybe for later)
```{r}

regfit.full=regsubsets(target~., data=crime_data, nvmax = 13)
reg.summary = summary(regfit.full)
names(reg.summary)
plot(reg.summary$cp, xlab = "Number of Variables", ylab="Cp")
which.min(reg.summary$cp)
points(5,reg.summary$cp[5],pch=20,col="red")
plot(regfit.full,scale="Cp")
coef(regfit.full,5)

#we must rename the target column in order for the bestglm function to work
crime_data <- rename(crime_data,y=target)
summary(crime_data)
bm <- bestglm(crime_data, family = binomial(link = "logit"), IC = "AIC", qLevel = 0.99, TopModels = 5,
method = "exhaustive", intercept = TRUE, weights = NULL,
nvmax = "default", RequireFullEnumerationQ = FALSE)

bm.bic <- bestglm(crime_data, family = binomial(link = "logit"), IC = "BIC", qLevel = 0.99, TopModels = 5,
method = "exhaustive", intercept = TRUE, weights = NULL,
nvmax = "default", RequireFullEnumerationQ = FALSE)

bm$BestModel
bm.bic$BestModel
table(true = target, pred = round(fitted(glm.fit)))
table(true = target, pred = round(fitted(bm$BestModel)))
table(true = target, pred = round(fitted(bm.bic$BestModel)))
```

#ROC Curve analysis

```{r}
library(pROC)
auc(crime_data$y, glm.fit$fitted.values)
auc(crime_data$y, gmu$fitted.values)
auc(crime_data$y, gmu.bic$fitted.values)
```